---
title: "VLSM data preprocessing"
author: "Pia Siegel"
date: "2023-11-14"
output: html_document
---

# Packages
```{r}
# Load packages
library(here)  
library(tidyverse) 
library(writexl)
library(dplyr)
library(fitdistrplus)
```

# Load data
```{r}
# Load already filtered data frame (Drop outs: 4002, 3364, 2884, 2998, 3392, 3960, 3154, 3994, 4086, 2522)
load(here::here("./RData/df_subsample_n41.RData"))
data_raw <- df_subsample
```

# Additional drop out
```{r}
# Remove participant with ID = 4227 due to lack of understanding of the task (ALWAYS pressed correct (accuracy - incorrect: 0.0000000; accuracy - correct: 1.0000000))
data_raw <- data_raw[!data_raw$ID == 4227, ]
```

# Data cleaning and preparing data frame
```{r}
# Rename "ID" into "RegistryCode"
data_raw <- data_raw %>% 
  rename(RegistryCode = ID)

# Add "sub" (subject) to the RegistryCode (ID)
data_raw$RegistryCode <- paste("sub", data_raw$RegistryCode)
   
# Create as well as calculate variable 'Weeks since Onset (WSO)'
## Define variables 'test' and 'onset' as dates
data_raw$Test <- as.Date(data_raw$Test, format = "%d.%m.%y")
data_raw$Onset <- as.Date(data_raw$Onset, format = "%d.%m.%y")

## Calculate 'WSO' and define variable as a numeric data type
data_raw$WSO <- difftime(data_raw$Test, data_raw$Onset, units = "weeks")
data_raw$WSO <- as.numeric(data_raw$WSO)

# Remove NA trials (accuracy) --> participants were too slow to react on time or didn't react at all
data_raw <- filter(data_raw, !(is.na(accuracy)))

# Remove RTs before picture was shown (equals rt = 0) and RTs below 200 ms
data_raw <- data_raw[!(data_raw$response_time_absolute < data_raw$time_picture),]
data_raw <- data_raw[!(data_raw$response_time < 200),]

# Remove RTs that deviated more than 3 SD from the mean per participant, condition and run
outlier_df <- data_raw %>% 
  group_by(RegistryCode, condition, run) %>% 
  mutate(outlier = ifelse(response_time > (mean(response_time) + 3*sd(response_time)), 1, 0))

n_outlier <- outlier_df %>% 
 group_by(RegistryCode, condition, run)  %>% 
 summarise(sum = sum(outlier))  

data_raw <- data_raw[outlier_df$outlier == 0, ]

# Make numeric variable for accuracy
data_raw$accuracy_num <- ifelse(data_raw$accuracy == "correct", 1, ifelse(data_raw$accuracy == "incorrect", 0, NA))
```

# Create data frames for analyses with SVR-LSM GUI
## Accuracy 
### Create .csv-dataframe for accuracy (condition - WPM)
```{r}
# Create new data frame with relevant variables 
df_acc_WPM <- data_raw[c('RegistryCode',
                         'Age',
                         'WSO',
                         'run',
                         'Lesion_volume',
                         'condition',
                         'accuracy_num')]

# Calculate mean of accuracy per run for each participant
df_acc_WPM_summary <- df_acc_WPM %>% 
  group_by(RegistryCode, condition, run) %>% 
  mutate(average_accuracy = mean(accuracy_num)) %>% 
  dplyr::select(-accuracy_num) %>% 
  distinct()
 
# Define runs to exclude if participant performed at or below chance level in WPM trials
df_acc_WPM_summary <- df_acc_WPM_summary %>% 
  group_by(RegistryCode, condition, run) %>% 
  pivot_wider(names_from = condition, values_from = average_accuracy) %>% 
  filter(WPM > 0.5) %>% 
  group_by(RegistryCode) %>% 
  mutate(WPM = mean(WPM), FPM = mean(FPM)) %>% 
  dplyr::select(-run) %>% 
  distinct()

# Extract data frame "data_acc_WPM.csv"
write.csv(df_acc_WPM_summary, here("./RData/data_acc_WPM.csv"), row.names = F)
```

### Create .csv-dataframe for accuracy (condition - FPM)
```{r}
# Create new data frame with relevant variables 
df_acc_FPM <- data_raw[c('RegistryCode',
                         'Age',
                         'WSO',
                         'run',
                         'Lesion_volume',
                         'condition',
                         'accuracy_num')]

# Calculate mean of accuracy per run for each participant
df_acc_FPM_summary <- df_acc_FPM %>% 
  group_by(RegistryCode, condition, run) %>% 
  mutate(average_accuracy = mean(accuracy_num)) %>% 
  dplyr::select(-accuracy_num) %>% 
  distinct()
 
# Define runs to exclude if participant performed at or below chance level in FPM trials
df_acc_FPM_summary <- df_acc_FPM_summary %>% 
  group_by(RegistryCode, condition, run) %>% 
  pivot_wider(names_from = condition, values_from = average_accuracy) %>% 
  filter(FPM > 0.5) %>% 
  group_by(RegistryCode) %>% 
  mutate(FPM = mean(FPM), WPM = mean(WPM)) %>% 
  dplyr::select(-run) %>% 
  distinct()

# Extract data frame "data_acc_FPM.csv"
write.csv(df_acc_FPM_summary, here("./RData/data_acc_FPM.csv"), row.names = F)
```

### Create .csv-dataframe for accuracy (congruency - incongruent)
```{r}
# Create new data frame with relevant variables 
df_acc_incongr <- data_raw[c('RegistryCode',
                             'Age',
                             'WSO',
                             'run',
                             'Lesion_volume',
                             'congruency',
                             'accuracy_num')]

# Calculate mean of accuracy for each participant
df_acc_incongr_summary <- df_acc_incongr %>% 
  group_by(RegistryCode, congruency, run) %>% 
  mutate(average_accuracy = mean(accuracy_num)) %>% 
  dplyr::select(-accuracy_num) %>% 
  distinct()

# Define runs to exclude if participant performed at or below chance level in incongruent trials
df_acc_incongr_summary <- df_acc_incongr_summary %>% 
  group_by(RegistryCode, congruency, run) %>% 
  pivot_wider(names_from = congruency, values_from = average_accuracy) %>% 
  filter(incongruent > 0.5) %>% 
  group_by(RegistryCode) %>% 
  mutate(incongruent = mean(incongruent), congruent = mean(congruent)) %>% 
  dplyr::select(-run) %>% 
  distinct()

# Extract data frame "data_acc_incongr"
write.csv(df_acc_incongr_summary, here("./RData/data_acc_incongr.csv"), row.names = F)
```

## Reaction time
### Create .csv-dataframe for reaction time (condition - WPM)
```{r}
# Create data frame with relevant variables 
df_rt_WPM <- data_raw[c('RegistryCode',
                        'Age',
                        'WSO',
                        'run',
                        'Lesion_volume',
                        'condition',
                        'accuracy_num',
                        'response_time')]

# Only keep trials for WPM condition and exclude runs if participant performed at or below chance level
df_acc_WPM_rt <- df_rt_WPM %>% 
  filter(condition == "WPM") %>% 
  group_by(RegistryCode, run) %>%
  mutate(average_accuracy = mean(accuracy_num)) %>% 
  filter(average_accuracy > 0.5) %>% 
  dplyr::select(RegistryCode, run) %>% 
  distinct()

# Now use this df to filter rt df
df_rt_WPM <- left_join(df_acc_WPM_rt, df_rt_WPM, by = c("RegistryCode", "run"))

# Remove incorrect trials from data frame, then log-transform RT, and finally summarise mean RT per participant
df_rt_WPM_summary <- df_rt_WPM %>% 
  filter(accuracy_num == 1) %>% # remove incorrect trials
  mutate(response_time_log = log(response_time)) %>% # log-transform rt
  dplyr::select(-c(response_time, accuracy_num)) %>% # drop unused columns
  group_by(RegistryCode, condition) %>% 
  mutate(average_rt_log = mean(response_time_log)) %>% # calculate average log-RT per condition
  dplyr::select(-c(response_time_log, run)) %>% # drop unused columns
  distinct() %>% 
  pivot_wider(names_from = condition, values_from = average_rt_log) # bring df into wide format 

# Extract data frame "data_rt_WPM"
write.csv(df_rt_WPM_summary, here("./RData/data_rt_WPM.csv"), row.names = F)
```

### Create .csv-dataframe for reaction time (condition - FPM)
```{r}
# Create data frame with relevant variables 
df_rt_FPM <- data_raw[c('RegistryCode',
                        'Age',
                        'WSO',
                        'run',
                        'Lesion_volume',
                        'condition',
                        'accuracy_num',
                        'response_time')]

# Only keep trials for FPM condition and exclude runs if participant performed at or below chance level
df_acc_FPM_rt <- df_rt_FPM %>% 
  filter(condition == "FPM") %>% 
  group_by(RegistryCode, run) %>%
  mutate(average_accuracy = mean(accuracy_num)) %>% 
  filter(average_accuracy > 0.5) %>% 
  dplyr::select(RegistryCode, run) %>% 
  distinct()

# Now use this df to filter rt df
df_rt_FPM <- left_join(df_acc_FPM_rt, df_rt_FPM, by = c("RegistryCode", "run"))

# Remove incorrect trials from data frame, then log-transform RT, and finally summarise mean RT per participant
df_rt_FPM_summary <- df_rt_FPM %>% 
  filter(accuracy_num == 1) %>% # remove incorrect trials
  mutate(response_time_log = log(response_time)) %>% # log-transform rt
  dplyr::select(-c(response_time, accuracy_num)) %>% # drop unused columns
  group_by(RegistryCode, condition) %>% 
  mutate(average_rt_log = mean(response_time_log)) %>% # calculate average log-RT per condition
  dplyr::select(-c(response_time_log, run)) %>% # drop unused columns
  distinct() %>% 
  pivot_wider(names_from = condition, values_from = average_rt_log) # bring df into wide format 

# Extract data frame "data_rt_FPM"
write.csv(df_rt_FPM_summary, here("./RData/data_rt_FPM.csv"), row.names = F)
```

### Create .csv-dataframe for reaction time (congruency - incongruent)
```{r}
# Create data frame with relevant variables 
df_rt_incongr <- data_raw[c('RegistryCode',
                            'Age',
                            'WSO',
                            'run',
                            'Lesion_volume',
                            'congruency',
                            'accuracy_num',
                            'response_time')]

# Only keep incongruent trials and exclude runs if participant performed at or below chance level
df_acc_incongr_rt <- df_rt_incongr %>% 
  filter(congruency == "incongruent") %>% 
  group_by(RegistryCode, run) %>%
  mutate(average_accuracy = mean(accuracy_num)) %>% 
  filter(average_accuracy > 0.5) %>% 
  dplyr::select(RegistryCode, run) %>% 
  distinct()

# Now use this df to filter rt df
df_rt_incongr <- left_join(df_acc_incongr_rt, df_rt_incongr, by = c("RegistryCode", "run"))

# Remove incorrect trials from data frame, then log-transform RT, and finally summarise mean RT per participant
df_rt_incongr_summary <- df_rt_incongr %>% 
  filter(accuracy_num == 1) %>% # remove incorrect trials
  mutate(response_time_log = log(response_time)) %>% # log-transform rt
  dplyr::select(-c(response_time, accuracy_num)) %>% # drop unused columns
  group_by(RegistryCode, congruency) %>% 
  mutate(average_rt_log = mean(response_time_log)) %>% # calculate average log-RT per condition
  dplyr::select(-c(response_time_log, run)) %>% # drop unused columns
  distinct() %>% 
  pivot_wider(names_from = congruency, values_from = average_rt_log) # bring df into wide format 

# Extract data frame "data_rt_incongr"
write.csv(df_rt_incongr_summary, here("./RData/data_rt_incongr.csv"), row.names = F)
```
